---
title: "WorkFlow"
author: "GenDoc94"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(writexl)
library(gridExtra)
library(patchwork)
library(ggplot2)
suppressPackageStartupMessages({
  library(tidyverse)
  library(readxl)
  library(jsonlite)
})
start_time <- Sys.time()
```

```{r starting_time, echo=FALSE, message=TRUE}
cat("El análisis comenzó a las:", as.character(start_time))
```

## OGM Read Workflow

Firstly, we have to pull all archives from our server (Dropbox) to our repository (OGMRead).

We should copy:

-   json_files

-   variants_files

-   aneuploidy_files

-   clinical_files

```{r copying_files, echo=FALSE, message=TRUE}
source("scripts/copy_files.R")
```

If there is no error here, all is OK (no replicate Ids) we can proceed to the next step.

## Making the bases

Now, we are going to run the code to create the complete base.

```{r reading_files, echo=FALSE, message=TRUE, warning=FALSE}
source("scripts/read_files.R")
```

After this we have some bases:

-   `complete_base` -\> All the information about the patient and clinics

-   `cb_filter` -\> All the important variants of patients without LB/B variants and with some standard parameters: VAF \> 0.05 and molecule count \> 10.

And some subbases:

-   `demograph`

-   `metadata`

-   `lecturas`

-   `quality`

-   `variants`

-   `aneuploidies`

## Basics of our data

```{r basics_data, echo=FALSE}
cat("Total muestras procesadas: ", total_samples, "\n")
cat("Total muestras analizadas: ", nrow(cb_filter), "\n")
cat("Total muestras pendientes: ", pending_samples, "\n")
cat("Total muestras fallidas: ", error_samples, "\n")
cat("Porcentaje muestras fallidas: ", round(error_samples/total_samples, 2)*100, "%")
```

```{r plot_evolucion_detallada, echo=FALSE, message=FALSE, warning=FALSE}
library(lubridate)

# 1. Preparar los datos con los 3 estados originales
evolucion_detallada <- metadata %>%
  mutate(
    # Agrupamos por mes (o "week" para más detalle)
    Mes = floor_date(as.Date(Fecha), "month"),
    # Convertimos el número a texto descriptivo para la leyenda
    Estado_Texto = case_when(
      Estado_Muestra == 3 ~ "Finalizado",
      Estado_Muestra == 2 ~ "Pendiente",
      Estado_Muestra == 1 ~ "Perdida",
      TRUE ~ "Desconocido"
    )
  ) %>%
  group_by(Mes, Estado_Texto) %>%
  summarise(Total = n(), .groups = "drop")

# 2. Gráfica con tres líneas
ggplot(evolucion_detallada, aes(x = Mes, y = Total, color = Estado_Texto, group = Estado_Texto)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2.5) +
  # Colores específicos: Verde (OK), Naranja (Espera), Rojo (Error)
  scale_color_manual(values = c(
    "Finalizado" = "#2E7D32", 
    "Pendiente"  = "#F9A825", 
    "Perdida"    = "#C62828"
  )) +
  labs(
    title = "Evolución Temporal del Estado de las Muestras",
    subtitle = "Seguimiento por mes: Finalizadas, Pendientes y Perdidas",
    x = "Fecha",
    y = "Número de Muestras",
    color = "Estado"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )
```

```{r viscosity}
visco_quanti <- metadata |> left_join(lecturas, by = c("Id" = "NumBN_L")) |>
        select(Id, Visco_grado, Media_Lectura)

p1 <- metadata |> 
  filter(!is.na(Visco_grado)) |> 
  mutate(Visco_grado = as.factor(Visco_grado)) |> 
  ggplot(aes(x = Visco_grado, fill = Visco_grado)) +
  geom_bar() +
  # Añadir los números encima
  stat_count(geom = "text", 
             aes(label = after_stat(count)),
             vjust = -0.5,      # Empuja el texto hacia arriba (valor negativo)
             size = 4,          # Tamaño de la fuente
             fontface = "bold") +
  theme_minimal() +
  labs(
    title = "Distribución del grado de viscosidad",
    x = "Grado de Viscosidad",
    y = "Número de muestras",
    fill = "Grado"
  ) +
  theme(legend.position = "none") +
  # Expandir el eje Y un poco para que el número no se corte arriba
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)))


p2 <- visco_quanti |> 
  filter(!is.na(Visco_grado), !is.na(Media_Lectura)) |> 
  mutate(Visco_grado = as.factor(Visco_grado)) |> 
  ggplot(aes(x = Visco_grado, y = Media_Lectura, fill = Visco_grado)) +
# Gráfico de Violín
  geom_violin(alpha = 0.6, trim = TRUE) + 
  geom_hline(yintercept = c(30, 120), color = "gold", linetype = "dashed", linewidth = 0.8) +
  geom_hline(yintercept = c(45, 90), color = "forestgreen", linetype = "solid", linewidth = 0.8) +
  theme_minimal() +
  labs(
    title = "[DNA] por grado de viscosidad",
    x = "Grado de Viscosidad",
    y = "[DNA] (ng/uL)",
    fill = "Grado"
  ) +
  theme(legend.position = "none")

p1 + p2
```

## Quality of Samples

Here...

```{r quality, echo=FALSE}
#Creating base
#Merge metadata+quality
qualmet <- left_join(quality, metadata, by = c("samplename" = "Id"))
qualmet_clean <- qualmet |> select(quantity, mol_n_50, coverage, map_rate, fp_rate, fn_rate, Dx)
```

### Total Sample
```{r total_quality, echo=FALSE}
cat("n=", nrow(qualmet_clean),"\n")
qualmet_clean |> select(-Dx) |> summary()
```

### MM
```{r MM_quality, echo=FALSE}
qualmet_clean_MM <- qualmet_clean |> filter(Dx == "MM") 
cat("n =", nrow(qualmet_clean_MM), "\n")
qualmet_clean_MM |> select(-Dx) |> summary()
```

### LMA
```{r LMA_quality, echo=FALSE}
qualmet_clean_LMA <- qualmet_clean |> filter(Dx == "LMA") 
cat("n =", nrow(qualmet_clean_LMA), "\n")
qualmet_clean_LMA |> select(-Dx) |> summary()
```

### LLC
```{r LLC_quality, echo=FALSE}
qualmet_clean_LLC <- qualmet_clean |> filter(Dx == "LLC") 
cat("n =", nrow(qualmet_clean_LLC), "\n")
qualmet_clean_LLC |> select(-Dx) |> summary()
```


## Generating list to export

```{r generating_export_list}
export_base <- complete_base |> select(Id, NHC, FechaNac, FechaPet, FechaDx, Petic, Dx)

#MM
export_base_MM <- export_base |> filter(Dx == "MM")
write_xlsx(export_base_MM, path = "bases/export_base_MM.xlsx")
saveRDS(export_base_MM, "export_base_MM.rds")
#envir = new.env() just to make sure start_time does not desapear
rmarkdown::render("export_base_MM.Rmd", output_dir = "bases", envir = new.env())
file.remove("export_base_MM.rds")

#LMA
export_base_LMA <- export_base |> filter(Dx == "LMA")
write_xlsx(export_base_LMA, path = "bases/export_base_LMA.xlsx")
saveRDS(export_base_LMA, "export_base_LMA.rds")
rmarkdown::render("export_base_LMA.Rmd", output_dir = "bases", envir = new.env())
file.remove("export_base_LMA.rds")

#LLA
export_base_LLA <- export_base |> filter(Dx == "LAL-B")
write_xlsx(export_base_LLA, path = "bases/export_base_LLA.xlsx")
saveRDS(export_base_LLA, "export_base_LLA.rds")
rmarkdown::render("export_base_LLA.Rmd", output_dir = "bases", envir = new.env())
file.remove("export_base_LLA.rds")
```

## Generating Graphs

Finally, we can generate our circles plot:

```{r generating_circleplots}
source("scripts/circleplot.R") #This source generate the function circleplot(id_pac, dbase)
source("scripts/pdfgraphs.R") #This source generate the function make_pdf_graphs(filter_dx, filename)

# Generate PDF for MM
#make_pdf_graphs(filter_dx = "MM", filename = "graphs/MM_graphs.pdf")

# Generate PDF for LMA
#make_pdf_graphs(filter_dx = "LMA", filename = "graphs/LMA_graphs.pdf")

# Generate PDF for LLC
#make_pdf_graphs(filter_dx = "LLC", filename = "graphs/LLC_graphs.pdf")
```

## Saving data

```{r saving_data, echo=FALSE, message=TRUE}
objects_ogm <- list(
  complete_base = complete_base,
  cb_filter = cb_filter
)

saveRDS(objects_ogm, "workflow_data.rds")
message("Saved: ", paste(names(objects_ogm), collapse = ", "))
```



```{r ending_time, cache = FALSE, echo=FALSE, message=TRUE}
end_time <- Sys.time()
cat("El análisis terminó a las:", as.character(end_time),"\n")
total_time <- (end_time - start_time)
cat("Ha tardado:", round(as.numeric(total_time),2), units(total_time))
```